---
title: "Parameter Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parameter Selection}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

The smoothLRC method uses a smoothing parameter ($/lambda$) and a rank parameter ($k$) to produce results. This article demonstrates the selection of these parameters.

```{r Setup, message=FALSE, warning=FALSE, collapse=TRUE}
library(smoothLRC)
library(ggplot2)
library(dplyr)
set.seed(1)
```

## Load Data

Using the [spatialLIBD](http://spatial.libd.org/spatialLIBD/) package, we'll load sample 151671 from the DLPFC data. 

```{r Load Data, message=FALSE, warning=FALSE}
sample <- "151671"

sce <- spatialLIBD::fetch_data(type = "sce")
sce <- sce[, sce$sample_name == sample]
```

## Parameter Grid

Here we'll consider $/lambda=(5, 10, 15)$ and $k=(10, 20)$. Further, we'll use 100 pixels as a test size.

```{r Parameters, message=FALSE, warning=FALSE}
lambdas <- c(5, 10, 15)
ks <- c(10, 20)
parameter_grid <- expand.grid(lambda = lambdas, k = ks)
test_size <- 100
```


## Cross-Validation

Next, we'll use cross-validation to select the optimal parameter values.

```{r smoothLRC, fig.height=5, fig.width=5.5, message=FALSE, warning=FALSE}
metrics <- NULL
for(combo in 1:nrow(parameter_grid)){
  print(parameter_grid[combo,])
  combination <- smooth_cv(sce, lambda = parameter_grid[combo,"lambda"], k = parameter_grid[combo,"k"], test_size, maxiter = 5)
  metrics <- rbind(metrics, tibble(lambda = parameter_grid[combo,"lambda"], 
                                   k = parameter_grid[combo,"k"],
                                   penal_lik = combination$penal_like,
                                   cv_like = combination$cv_like)
                   )
}

metrics %>% 
  ggplot(aes(x = lambda, y = cv_like)) +
  geom_point() +
  geom_line() +
  facet_wrap(~k) +
  ylab("CV Likelihood")

metrics %>% 
  group_by(k) %>%
  slice_max(order_by = cv_like, n = 1) %>%
  ungroup() %>%
  mutate(Selected = ifelse(penal_lik == max(penal_lik), "Yes", "No")) %>%
  ggplot(aes(x = k, y = penal_lik, label = lambda)) +
  geom_point() +
  geom_line() +
  geom_label(aes(color = Selected)) +
  ylab("Penalized Likelihood")

```


We can see that using the recommend procedure, the selected combination is ($\lambda=15$, $k=20$). In practice, we recommend running for a much larger number of iterations (e.g. 1,000) and looking over a larger parameter space. 
